---
title: "Fuel consumption versus transmission type"
author: "Ilya Maltsev"
date: '23 Apr 2017'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

Fuel consumption seems to be one of the most important aspects of car usage. 
Usually the less fuel is expended by a car the more usable the car is.

The study is focused on relationship between fuel consumption (measured as 
miles per gallon, MPG) and transmission type (automatic or manual).

The main questions are:

1. Is an automatic or manual transmission better for MPG?

2. Quantify the MPG difference between automatic and manual transmissions.

The data is used from _Motor Trend_ US magazine data.

## Exploratory data analysis

### About the data

The *mtcars* dataset from *datasets* package is used:

1. *mpg* - Miles/(US) gallon.
2. *am* - transmission (0 = automatic, 1 = manual).

```{r, dataset}
require(datasets)
head(mtcars, 3)
```

The dataset contains `r nrow(mtcars)` observations.

## Modelling

## Conclusions
