---
title: "Fuel consumption versus transmission type"
author: "Ilya Maltsev"
date: '23 Apr 2017'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

Fuel consumption seems to be one of the most important aspects of car usage. 
Usually the less fuel is expended by a car the more usable the car is.

The study is focused on relationship between fuel consumption (measured as 
miles per gallon, MPG) and transmission type (automatic or manual).

The main questions are:

1. Is an automatic or manual transmission better for MPG?

2. Quantify the MPG difference between automatic and manual transmissions.

The data is used from _Motor Trend_ US magazine data.

## Exploratory data analysis

```{r, attach.libraries, message=F, warning=F}
require(datasets)
require(ggplot2)
require(corrplot)
```

### About the data

The *mtcars* dataset from *datasets* package is used.

```{r, dataset}
data("mtcars")
head(mtcars, 3)
```

The dataset contains `r nrow(mtcars)` observations.

### *MPG* distribution

The *mpg* data does not follow normal distribution (see Appendix). This means 
that Students t-test cannot be used to get statistical evidence of *mpg* 
to *am* dependency.

### Correlation matrix

Some of the *mtcars* variables are physically related, e.g. displacement and 
horse power. Ideally, linear models should include only independent variables. 

```{r, correlation.matrix, fig.width=4, fig.height=4}
corrmatrix = cor(mtcars)
corrplot(corrmatrix, method = "circle", order ="FPC")
```

Correlation matrix shows that the *mpg* correlates with the *cyl*, 
*disp*, *hp*, *wt* the most. However the *cyl*, *disp*, *hp* are highly
correlated with each other. The *wt* correlates with the *hp* less than with 
the *cyl* and *disp*.

## Modelling

Consider the linear model $mpg=wt+hp+am$ according to the correlation matrix.

```{r, fit.mpg_wt_hp_am}
fit.mpg_wt_hp_am = lm(formula = mpg ~ wt + hp + I(factor(am)), data = mtcars)
round(summary(fit.mpg_wt_hp_am)$coef, 4)
```

The value for *am* in the last column shows that the we failed to reject 
the null hypothesis: the *am* does not affect the *mpg*. The analysis of 
variance table (see Appendix) confirms this result.

## Conclusions

The *mtcars* data analysis does not confirm that there is a causal relationship 
between fuel consumption and transmission type.

## Appendix

### Normality of the *MPG* data

[Shapiro-Wilk test](https://en.wikipedia.org/wiki/Shapiro%E2%80%93Wilk_test) is
used to verify the normality of the *mpg* data. Level 
of statistical significance is taken to be 0.05. Results for the *mpg* and with
distinction by cylinders count or by transmission type are shown below:

```{r, mpg.normality.shapiro.wilk}
rbind(
    shapiro.test(mtcars$mpg[mtcars$cyl == 4]),
    shapiro.test(mtcars$mpg[mtcars$cyl == 6]),
    shapiro.test(mtcars$mpg[mtcars$cyl == 8]),
    shapiro.test(mtcars$mpg[mtcars$am == 0]),
    shapiro.test(mtcars$mpg[mtcars$am == 1]),
    shapiro.test(mtcars$mpg)
)[, c("data.name", "p.value")]
```

All the p-values are greater than 0.05 so we reject the null hypothesis that 
the data are normal.

### Analysis of variance table

```{r, anova}
fit.mpg_wt = lm(formula = mpg ~ wt, data = mtcars)
fit.mpg_wt_hp = lm(formula = mpg ~ wt + hp, data = mtcars)
anova(fit.mpg_wt, fit.mpg_wt_hp, fit.mpg_wt_hp_am)
```

The analysis shows that the *wt* and the *hp* are important for predicting the 
*mpg* and the *am* has low importance.
