---
title: "GLM_Exercises"
author: "Ilya Maltsev"
date: '20 Apr 2017'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, libraries}
library(dplyr)
```

## Binary generalized linear regression

### Question 1

Load the dataset Seatbelts as part of the datasets package via data(Seatbelts). Use
as.data.frame to convert the object to a dataframe. Create a new outcome variable for
whether or not greater than 119 drivers were killed that month. Fit a logistic regression GLM with this variable as the outcome and kms, PetrolPrice and law as predictors. Interpret your parameters.

```{r, q1prepare}
data("Seatbelts")
seatbelts = as.data.frame(Seatbelts)
seatbelts = seatbelts %>% 
    mutate(mmc = (kms - mean(kms))/1000) %>%
    mutate(pp = (PetrolPrice - mean(PetrolPrice))/sd(PetrolPrice)) %>%
    mutate(extra = DriversKilled > 119)
```

```{r, q1fit}
q1fit = glm(data = seatbelts, 
            formula = extra ~ mmc + pp + law, 
            family = binomial)
summary(q1fit)
```

Distance has low effect on a count of days with extra amount of drivers killed:

```{r, q1interpretation1}
print(q1fit$coef[2])
print(exp(q1fit$coef[2]))
```

Increasing petrol price and law have an effect on decreasing of days count 
with extra amount of drivers killed:

```{r, q1interpretation2}
print(q1fit$coef[3])
print(exp(q1fit$coef[3]))
print(q1fit$coef[4])
print(exp(q1fit$coef[4]))
```

Only petrol price decreasing has statistical significance because only its 
confidence interval does not include 1:
```{r, q1confint}
print(exp(confint(q1fit)))
```

### Question 2

Fit a binomial model with DriversKilled as the outcome and drivers as the total 
count with kms, PetrolPrice and law as predictors, interpret your results.

```{r, q2prepare}
data("Seatbelts")
seatbelts = as.data.frame(Seatbelts)
seatbelts = seatbelts %>% 
    mutate(mmc = (kms - mean(kms))/1000) %>%
    mutate(pp = (PetrolPrice - mean(PetrolPrice))/sd(PetrolPrice))
```

```{r, q2fit}
q2fit = glm(data = seatbelts, 
            formula = cbind(DriversKilled, drivers - DriversKilled) ~ mmc + pp + law, 
            family = binomial)
summary(q2fit)
```

```{r, q2interpretation}
print(q2fit$coef[2])
print(exp(q2fit$coef[2]))
print(q2fit$coef[3])
print(exp(q2fit$coef[3]))
print(q2fit$coef[4])
print(exp(q2fit$coef[4]))
print(exp(confint(q2fit)))
```

### Question 3

Refer to Question 1. Use the anova function to compare models with just law, law and
PetrolPrice and all three predictors.

```{r, q3prepare}
data("Seatbelts")
seatbelts = as.data.frame(Seatbelts)
seatbelts = seatbelts %>% 
    mutate(mmc = (kms - mean(kms))/1000) %>%
    mutate(pp = (PetrolPrice - mean(PetrolPrice))/sd(PetrolPrice)) %>%
    mutate(extra = DriversKilled > 119)
```

```{r, q3fit}
q31fit = glm(data = seatbelts, 
             formula = extra ~ law, 
             family = binomial)
q32fit = glm(data = seatbelts, 
             formula = extra ~ law + pp, 
             family = binomial)
q33fit = glm(data = seatbelts, 
             formula = extra ~ law + pp + mmc, 
             family = binomial)
anova(q31fit, q32fit, q33fit)
```
