---
title: "GLM_Exercises2"
author: "Ilya Maltsev"
date: '20 Apr 2017'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, libraries}
library(dplyr)
library(ggplot2)
```

## Poisson model

### Question 1

Load the dataset Seatbelts as part of the datasets package via data(Seatbelts). 
Use as.data.frame to convert the object to a dataframe. Fit a Poisson 
regression GLM with UKDriversKilled as the outcome and kms, PetrolPrice and 
law as predictors. Interpret your results.

```{r, q1prepare}
data("Seatbelts")
seatbelts = as.data.frame(Seatbelts)
seatbelts = seatbelts %>% 
    mutate(pp = (PetrolPrice - mean(PetrolPrice))/sd(PetrolPrice)) %>%
    mutate(mmc = (kms - mean(kms))/1000)
```

```{r, q1fit}
q1fit = glm(data = seatbelts, 
            formula = DriversKilled ~ pp + law + mmc, 
            family = poisson)
summary(q1fit)
```

Increasing of normalized petrol price per unit causes decreasing of expected 
drivers killed on `r round((1 - exp(q1fit$coef[2])) * 100, 2)`%.

Introducing law causes decreasing of expected 
drivers killed on `r round((1 - exp(q1fit$coef[3])) * 100, 2)`%.

Increasing of distance per unit causes decreasing of expected 
drivers killed on `r round((1 - exp(q1fit$coef[4])) * 100, 2)`%.

### Question 2

Refer to question 1. Fit a linear model with the log of drivers killed as 
the outcome. Interpret your results.

```{r, q2fit}
q2fit = lm(data = seatbelts, 
           formula = I(log(DriversKilled)) ~ pp + law + mmc)
summary(q2fit)
```

### Question 3

Refer to question 1. Fit your Poisson log-linear model with drivers as a 
log offset (to consider the proportion of drivers killed of those killed 
or seriously injured.)

```{r, q3fit}
q3fit = glm(data = seatbelts, 
            formula = DriversKilled ~ pp + law + mmc,
            family = poisson,
            offset = log(drivers))
summary(q3fit)
```

### Question 4

Refer to Question 1. Use the anova function to compare models with just law, 
law and PetrolPrice and all three predictors.

```{r, q4fit}
q41fit = glm(data = seatbelts, 
            formula = DriversKilled ~ law, 
            family = poisson)
q42fit = update(q41fit, 
                formula = DriversKilled ~ law + pp)
q43fit = update(q42fit, 
                formula = DriversKilled ~ law + pp + mmc)
anova(q41fit, q42fit, q43fit)
```
